---
title: "RepDataProj1"
author: "AllenKennedy"
date: "Wednesday, June 03, 2015"
output: 
  html_document:
    pandoc_args: ["+RTS", "-K64m","-RTS"]
---

##Loading required packages

```{r, results='hide'}
  library(dplyr)
  library(ggplot2)
  library(lattice)
```
##Loading the data
  
```{r}
setwd("C:/Users/kennedyal/RepData_PeerAssessment1")
activity<- read.csv("activity.csv")
```
##Analysis of the data file
```{r}
str(activity)
#As we can see the date is a factor, but we will keep it as a factor for our analysis
summary(activity)
#there are 2304 NA steps
```

##What is mean total number of steps taken per day?

###total number of steps per day
```{r}
day<- group_by(activity, date)
summary <-summarize(day, sum_steps = sum(steps, na.rm = TRUE))
summary

```

###Histogram of total number of steps each day

```{r}
hist(summary$sum_steps, col = "blue", xlab = "Steps", main = "Sum of Steps")
```

###Mean of total number of steps each day
```{r}
mean(summary$sum_steps)
```

###Median of total number of steps each day
```{r}
median(summary$sum_steps)
```

##What is the average daily activity pattern?

###Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
interval <- group_by(activity, interval)
interval_summary <-summarize(interval, avg_steps = mean(steps, na.rm = TRUE))
qplot(interval, avg_steps, data = interval_summary, geom = "line")
```
###Max 5 minute interval
```{r}
max<-which.max(interval_summary$avg_steps)
interval_summary[max,]
```

##Impute missing values
###count number of na rows
```{r}
cc <-complete.cases(activity)
a<- activity[!cc,]
nrow(a)
```
##impute missing values
```{r}
activity_impute <-activity
for (i in 1:nrow(activity_impute))
{
  if (is.na(activity_impute$steps[i]))
  {
    int <- activity_impute$interval[i]
    row <- which(interval_summary$interval == int)
    steps <- interval_summary$avg_steps[row]
    activity_impute$steps[i] <- steps
  }
  
}
```
##creating new data set grouped by date for total number of steps each day and plot Histogram for it
```{r}
day_impute<- group_by(activity_impute, date)
summary_imputed <-summarize(day_impute, sum_steps = sum(steps, na.rm = TRUE))
hist(summary_imputed$sum_steps, col = "blue", xlab = "Steps", main = "Sum of Steps")
```
###Report mean and median for before imputing and after imputing
```{r}
mean(summary$sum_steps)
median(summary$sum_steps)
mean(summary_imputed$sum_steps)
median(summary_imputed$sum_steps)
##As we can see there is a slight change after the NA were imputed
```
##Are there differences in activity patterns between weekdays and weekends?
### new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
activity_impute$date <- as.Date(activity_impute$date, "%Y-%m-%d")
activity_impute$day<- weekdays(activity_impute$date)
activity_impute <- mutate(activity_impute, weekend = factor(day %in% c("Saturday","Sunday"), labels = c("weekday","weekend")))
head(activity_impute)
```
###Time series plot
```{r}
week_agg <- aggregate(activity_impute$steps, by = list(activity_impute$interval, activity_impute$weekend),mean )
names(week_agg) <-c("interval","weekend","avg_steps")
xyplot(avg_steps ~ interval | weekend,data = week_agg,type = "l", layout = c(1,2), ylab = "Number of steps")
```

